---
title: "Dissertation"
author: "Thomas Bridier"
date: "2024-03-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library and datasets 

```{r}
library(dplyr)
library(ggplot2)
library(esquisse)
library(readxl)
library(tidyr)
library(car)
library(lsmeans)
library(openxlsx)
library(corrplot)
library(stats)

```

```{r}

```

```{r}

```

```{df1 <- read_excel("C:/Users/thomb/OneDrive - University of Edinburgh/4th year/dissertation project/bibliography/dataset/Diss project/excel file/actual_dataset.xlsx",}
    sheet = "record_sheet")
df1 <- df1[df1$site_status != "Not Available" & df1$ISO3 != "SWE", ]
columns_to_exclude <- c("7AC", "10", "11", "12", "13", "14", "15", "16", "18", "27", "28", "29", "30", "31AC")
df1.1 <- df1[, setdiff(names(df1), columns_to_exclude)]
columns_to_convert <- 34:56
# Convert the specified columns to numeric
df1.2 <- df1.1 %>%
  mutate_at(vars(columns_to_convert), as.numeric)

df2 <- pivot_longer(df1.2, cols = 34:55, names_to = "question", values_to = "score")
df2 <- df2 %>%
  mutate(category = case_when(
    question %in% c(1, 5, 6, 7, 34) ~ "A",
    question %in% c(2, 3, 4) ~ "B",
    question %in% c(8, 19, 20, 21, 22, 23, 24, 32, 33) ~ "C",
    question %in% c(9, 17, 25, 26, 31) ~ "D",
    TRUE ~ NA_character_  # Handle other cases if any
  ))
```

# RQ1 : How Effective is the UK offshore Protected Area Management? 

## What is the overall distribution of UK offshore PA management effectiveness ? 

### Barplot : 

```{r}
ggplot(df1.2) +
  aes(x = ratio) +
  geom_histogram(bins = 30L, fill = "#CCCDCD", color = "black", size = 0.5) +
  geom_vline(xintercept = 37.67404, color = "black", linetype = "dashed", size = 1) +  # Add vertical line
  labs(x = "Effectiveness score (%)", y = "Number of Assessments") +
  theme_bw(base_size = 15) +theme(panel.grid = element_blank(),axis.title = element_text(face = "bold"))  # Remove grid
```

### **Quartile calculation:** 

```{r}
mean_value <- mean(df1.2$ratio)
mean_value
percentage_below_33 <- sum(df1.2$ratio < 33)
percentage_below_67 <- sum(df1.2$ratio > 67)
percentage_between <- nrow(df1.2) - percentage_below_33 - percentage_below_67
cat("Number of values above 67%:", percentage_below_67, "\n")
cat("Number of values below 33%:", percentage_below_33, "\n")
cat("Number of values between 33% and 67%:", percentage_between, "\n")
```

## How does management effectiveness relate to PA characteristics ?

### **Size :** 

```{r}
lm1<-lm(log(ratio)~log(area_ha),data=df1.2)
summary(lm1)
plot(lm1)
ggplot(df1.2) +
  aes(x =log(ratio) , y = log(area_ha)) +
  geom_point(shape = "circle", size = 1.4, colour = "black") +  
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  # Add a linear trendline in blue
  theme_bw(15) +  # Apply a black and white theme
  theme(panel.grid = element_blank(),axis.title = element_text(face = "bold"))+labs(x = " log Protected area size (ha) ", y = "log effectiveness score (%)")
```

### Year the PA was established in : 

```{r}
lm3<-lm(ratio~STATUS_YR,data=df1.2)
summary(lm3)
plot(lm3)
ggplot(df1.2) +aes(x = STATUS_YR, y = ratio) + geom_point(shape = "circle", size = 2, colour = "black") + geom_smooth(method = "lm", se = TRUE, color = "grey") + theme_bw(15) + theme(panel.grid = element_blank(),axis.title = element_text(face = "bold"))+labs(x = "Year The Protected Area was established", y = "effectiveness score (%)")
```

### site status

```{r}
mod2 <- aov(log(ratio) ~ site_status, data = df1.2)
summary(mod2)
plot(mod2)
```

### distance from the shore 

```{r}
df1.2 <- df1.2 %>%
  mutate(NEAR_DIST_M = as.numeric(NEAR_DIST_M)) %>%
  filter(!is.na(NEAR_DIST_M))
lm7<-lm(log(ratio)~NEAR_DIST_M,data=df1.2)
summary(lm7)
```

## which aspects of management impacts effectiveness the most ?

###  Categories of management 

```{r}
df4 <- pivot_longer(df1.2, cols =60:63 , names_to = "Categories", values_to = "Ratio_cat")
df4 <- df4 %>%
  mutate(Categories = gsub("ratio_A", "A", Categories)) %>%
  mutate(Categories = gsub("ratio_B", "B", Categories)) %>%
  mutate(Categories = gsub("ratio_C", "C", Categories)) %>%
  mutate(Categories = gsub("ratio_D", "D", Categories))
lm4<-aov(Ratio_cat~Categories,data=df4)
summary(lm4)
posthoc <- TukeyHSD(lm4)
# Print the results
print(posthoc)
ggplot(df4) +
  aes(x = Categories, y = Ratio_cat,fill = Categories) +
  geom_boxplot() + theme_bw(15)+theme(panel.grid = element_blank(),axis.title = element_text(face = "bold"))+labs(x = "Categories of management", y = "Effectiveness score (%)")
```

### contribution score 

```{r}
df2 <- df2 %>%
  mutate(contribution = (score - average_score)*100) %>%
  ungroup()

mod1<-aov(contribution~question,data=df2)
summary(mod1)
posthoc <- lsmeans(mod1, "question", adjust = "tukey")
# View post hoc comparisons
posthoc
plot(posthoc)

df2$question <- as.factor(df2$question)
df2 <- df2[!(df2$question %in% c("28", "29", "30")), ]
# Convert question to factor with levels sorted numerically
df2$question <- factor(df2$question, levels = unique(df2$question))
df2$question <- factor(df2$question, levels = unique(df2$question[order(df2$category)]))
ggplot(df2) +
  aes(x = question, y = contribution, fill = category) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +theme_bw(15) +
 theme(panel.grid = element_blank(),axis.title = element_text(face = "bold")) +
  labs(x = "Question ID", y = "Contribution (%)")
```

# RQ2: How PA management effectiveness relate to the extent of bottom fishing restrictions and effort ?

## How much of UK PAs are concerned with fishery restrictions ? 

```{r}
df3.2.5 <- read_excel("C:/Users/thomb/OneDrive - University of Edinburgh/4th year/dissertation project/bibliography/dataset/Diss project/excel file/actual_dataset.xlsx", 
    sheet = "JNCC_data ")
df3.2.5 <- df3.2.5[df3.2.5$`Area (ha)` >= 7500, ]
df3.2.5 <- df3.2.5[df3.2.5$Site_name != "Dolphin Head", ]

mod3.2.5<-aov(score~`fishing_restriction`,data=df3.2.5)
summary(mod3.2.5)
posthoc <- TukeyHSD(mod3.2.5)

# Print the results
print(posthoc)

ggplot(df3.2.5) +
  aes(x = `fishing_restriction`, y = score) +
  geom_boxplot(fill = "#CCCDCD") + labs()+ # Neutral grey color for the box
  theme_bw(15) +  # Black and white theme
  theme(panel.grid = element_blank(),axis.title = element_text(face = "bold")) +
  # Change the X axis title
  xlab("Bottom Fishing Restriction") +ylab("effectiveness score (%)")+
  # Set Y axis limits
  ylim(0, 100)
```

## How does the UK PA management relate to fishing hours? 

### Total fishing effort - 

```{r}
fish_EU <- read_excel("C:/Users/thomb/OneDrive - University of Edinburgh/4th year/dissertation project/bibliography/dataset/Diss project/excel file/actual_dataset.xlsx", 
    sheet = "Fishing activity (2018-2021)")
fish_EU1 <- fish_EU[fish_EU$AREA_HA >= 7500, ]
lm3<-lm(`score PA`~`Total fishing / ha`,data=fish_EU1)
summary(lm3)
plot(lm3)

ggplot(fish_EU1) +
  aes(x = `score PA`, y = `Total fishing / ha`) +geom_point(shape = "circle", size = 2, colour = "black") +  # Change point size to 15 and color to black
  geom_smooth(method = "lm", se = TRUE, color = "grey") +  # Add a linear trendline in blue
  theme_bw(15) +  # Apply a black and white theme
  theme(panel.grid = element_blank(),axis.title = element_text(face = "bold"))+labs(x = "effectiveness score (%)", y = "Total fishing effort / ha")+ylim(-0.01,0.125)
```

### per geartypes - 

```{r}
fish_EU1$OT_fishing_per_ha <- fish_EU1$OT_fishing / fish_EU1$AREA_HA
fish_EU1$Beam_fishing_per_ha <- fish_EU1$Beam_fishing / fish_EU1$AREA_HA
fish_EU1$Seine_fishing_per_ha <- fish_EU1$Seine_fishing / fish_EU1$AREA_HA
fish_EU1$Dredge_fishing_per_ha <- fish_EU1$Dredge_fishing / fish_EU1$AREA_HA
lm3.1<-lm(`score PA`~OT_fishing_per_ha,data=fish_EU1)
lm3.2<-lm(`score PA`~Beam_fishing_per_ha,data=fish_EU1)
lm3.3<-lm(`score PA`~Seine_fishing_per_ha,data=fish_EU1)
lm3.4<-lm(`score PA`~Dredge_fishing_per_ha,data=fish_EU1)
summary(lm3.1)
summary(lm3.2)
summary(lm3.3)
summary(lm3.4)

# Graph for OT_fishing_per_ha
ggplot(fish_EU1) +
  aes(x = `score PA`, y = OT_fishing_per_ha) +
  geom_point(shape = "circle", size = 2, colour = "black") +
  geom_smooth(method = "lm", se = TRUE, color = "grey") +
  theme_bw(15) +
  theme(panel.grid = element_blank(),axis.title = element_text(face = "bold")) +
  labs(x = "effectiveness score (%)", y = "Otter fishing effort / ha")+ylim(-0.01, 0.125)

# Graph for Beam_fishing_per_ha
ggplot(fish_EU1) +
  aes(x = `score PA`, y = Beam_fishing_per_ha) +
  geom_point(shape = "circle", size = 2, colour = "black") +
  geom_smooth(method = "lm", se = TRUE, color = "grey") +
  theme_bw(15) +
  theme(panel.grid = element_blank(),axis.title = element_text(face = "bold")) +
  labs(x = "effectiveness score (%)", y = "Beam fishing effort / ha")+ylim(-0.01,0.125)

```

## How does PA fishing restrictions relate to fishing effort 

### Total 

```{r}
mod3.2.5<-aov(score~`fishing_restriction`,data=df3.2.5)
summary(mod3.2.5)
posthoc <- TukeyHSD(mod3.2.5)

# Print the results
print(posthoc)
ggplot(df3.2.5) +
  aes(x = fishing_restriction, y = `Total_fished _ha`) +
  geom_boxplot(fill = "grey80", color = "black") +  # Neutral grey color for the box
  theme_bw(15) +  # Black and white theme
  theme(panel.grid = element_blank(),axis.title = element_text(face = "bold")) +
  # Change the X axis title
  xlab("Bottom Fishing Restriction") +ylab("Total fishing effort / ha")
  # Set Y axis limits
```

### For BT and OT 

```{r}
mod3.2.6<-aov(BT_HA~fishing_restriction,data=df3.2.5)
summary(mod3.2.6)
mod3.2.6<-aov(OT_HA~fishing_restriction,data=df3.2.5)
summary(mod3.2.6)
```

# RQ3 : How does the UK perform  compared to Sweden ?

## How management effectiveness differs between Sweden and the UK 

```{r}
df7 <- read_excel("C:/Users/thomb/OneDrive - University of Edinburgh/4th year/dissertation project/bibliography/dataset/Diss project/excel file/actual_dataset.xlsx", 
    sheet = "record_sheet")
df7$site_status <- as.factor(df7$site_status)
df7 <- df7[df7$site_status != 'Not Available',]
df7<-df7[, -c(41, 44:50, 52,53,54, 61, 66)]
columns_to_convert <- 34:56
# Convert the specified columns to numeric
df7<- df7 %>%
  mutate_at(vars(columns_to_convert), as.numeric)
df7.2 <- pivot_longer(df7, cols = 34:56, names_to = "question", values_to = "score")
df7.3 <- pivot_longer(df7, cols =61:64 , names_to = "Categories", values_to = "Ratio_cat")

mod7<-aov(ratio ~ ISO3, data = df7)
summary(mod7)
plot(mod7)

ggplot(df7) +
  aes(x = ISO3, y = ratio, fill = ISO3) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +theme_bw(15)+theme(panel.grid = element_blank(),axis.title = element_text(face = "bold"))+labs(x = "Countries", y = " Effectiveness score (%)")+ scale_fill_manual(
    values = c(GBR = "#FFBFBF",
    SWE = "#BFBFFF"))
```

### per category - 

```{r}
df7.3<- df7.3 %>%
  mutate(Categories = gsub("ratio_A", "A", Categories)) %>%
  mutate(Categories = gsub("ratio_B", "B", Categories)) %>%
  mutate(Categories = gsub("ratio_C", "C", Categories)) %>%
  mutate(Categories = gsub("ratio_D", "D", Categories))

ggplot(df7.3) +
  aes(x = Categories, y = Ratio_cat, fill = ISO3) +
  geom_boxplot() + scale_fill_manual(
    values = c(GBR = "#FFBFBF",
    SWE = "#BFBFFF"))+
  theme_bw(15)+theme(panel.grid = element_blank(),axis.title = element_text(face = "bold"))+labs(x = "Management category", y = "Effectiveness score (%)")
mod3<-aov(Ratio_cat ~ ISO3*Categories, data = df7.3)
summary(mod3)

posthoc <- TukeyHSD(mod3)
# Print the results
print(posthoc)
plot(posthoc)
```

## UK & SWE Case study : The management of the most performing PAs 

```{r}
df_spider <- read_excel("C:/Users/thomb/OneDrive - University of Edinburgh/4th year/dissertation project/bibliography/dataset/Diss project/excel file/actual_dataset.xlsx", 
    sheet = "record_sheet")
columns_to_convert <- 34:69
# Convert the specified columns to numeric
df_spider <- df_spider %>%
  mutate_at(vars(columns_to_convert), as.numeric)
```

### Category A : 

```{r fig.height=6, fig.width=7}
df5.3 <- df1.2[df1.2$NAME == "Hatton Bank", ]

# Select specific columns from df5.2
df5.3 <- df5.3[, c("1", "5", "6", "7", "34")]

# Add a new row to df5.2 with the values corresponding to "Nordvästra Skånes havsområde"
new_row <- df_spider[df_spider$NAME == "Bratten", c("1", "5", "6", "7", "34")]
df5.3 <- rbind(rep(3, 5), rep(0, 5), df5.3, new_row)
colnames(df5.3) <- c("legal status","protected area design","PA boundary demarcation","management planning","connectivity")

# Assuming df5.2 has two rows represented
# Custom colors with transparency for each row
colors <- c(rgb(1, 0.5, 0.5, 0.5), rgb(0.5, 0.5, 1, 0.5))
  # Adjust the alpha value (0.5) for transparency

# Create the radar chart
radarchart(df5.3, axistype = 1, pcol = colors, pfcol = colors, plwd = 4, cglcol = "grey", cglty = 1, axislabcol = "black", cglwd = 0.8, vlcex = 1, cglcex = 1,axlabcex = 1)
```

### Category B : 

```{r fig.height=6, fig.width=7}
df5.4 <- df_spider[df_spider$NAME == "Hatton Bank", ]

# Select specific columns from df5.2
df5.4 <- df5.4[, c("2", "3", "4")]

# Add a new row to df5.2 with the values corresponding to "Nordvästra Skånes havsområde"
new_row <- df_spider[df_spider$NAME == "Bratten", c("2", "3", "4")]
df5.4 <- rbind(rep(3, 3), rep(0, 3), df5.4, new_row)
colnames(df5.4) <- c("protected area objectives","PA regulation / controls","Planning for adjacent land / sea use")
colors <- c(rgb(1, 0.5, 0.5, 0.5), rgb(0.5, 0.5, 1, 0.5))

# Create the radar chart
radarchart(
  df5.4,
  axistype = 1,
  pcol = colors,
  pfcol = colors,  # Use the custom colors with transparency for fill
  plwd = 4,
  cglcol = "grey",
   plcol = "black",
  cglty = 2,
  axislabcol = "black",
  cglwd = 0.8,
  vlcex = 1
)
```

```{r fig.height=6, fig.width=7}
# Assuming df1.2 is your original data frame

# Select rows from df1.2 with NAME equal to "The Canyons"
df5.2 <- df_spider[df_spider$NAME == "Hatton Bank", ]
# Select specific columns from df5.2
df5.2 <- df5.2[, c("8", "19", "20", "21", "22", "23", "24", "32", "33")]
# Add a new row to df5.2 with the values corresponding to "Bratten "
new_row <- df_spider[df_spider$NAME == "Bratten", c("8", "19", "20", "21", "22", "23", "24", "32", "33")]
df5.2 <- rbind(rep(3, 9), rep(0, 9), df5.2, new_row)
colnames(df5.2) <- c(
  "Regular Work Plan",
  "Research",
  "Monitoring and Evaluation",
  "Resource Management",
  "Climate Change",
  "Carbon Capture",
  "Ecosystem Services",
  "Livelihood Benefits",
  "Threats"
)
library(fmsb)

# Assuming df5.2 has two rows represented
# Custom colors with transparency for each row
new_colors <- c(rgb(1, 0.5, 0.5, 0.5), rgb(0.5, 0.5, 1, 0.5))
  # Adjust the alpha value (0.5) for transparency

# Custom names for the legend
legend_names <- c("The Canyons", "Nordvästra Skånes havsområde")

# Create the radar chart
radarchart(
  df5.2,
  axistype = 1,
  pcol = new_colors,
  pfcol = new_colors,  # Use the custom colors with transparency for fill
  plwd = 4,
  cglcol = "grey",
  cglty = 1,
  axislabcol = "black",
  cglwd = 0.8,
  vlcex = 0.8)
```

### Category D : 

```{r fig.height=6, fig.width=7}
df5.5 <- df_spider[df_spider$NAME == "Hatton Bank", ]

# Select specific columns from df5.2
df5.5 <- df5.5[, c("9", "17", "25","26","31")]

# Add a new row to df5.2 with the values corresponding to "Nordvästra Skånes havsområde"
new_row <- df_spider[df_spider$NAME == "Hatton Bank", c("9", "17", "25","26","31")]
df5.5 <- rbind(rep(3, 5), rep(0, 5), df5.5, new_row)

colnames(df5.5) <- c("resource inventory","protection systems","education and awareness","state and commercial neighbours","local communities")
radarchart(
  df5.5,
  axistype = 1,
  pcol = colors,
  pfcol = colors,  # Use the custom colors with transparency for fill
  plwd = 4,
  cglcol = "grey",
  cglty = 1,
  axislabcol = "black",
  cglwd = 0.8,
  vlcex = 0.8
)
```

# Appendix : 

## OSPAR management category 

```{r}
ggplot(df1.2) +
  aes(x = OSP_A) +
  geom_bar(fill = "grey80") + 
  labs(x = "Is the management documented ?", y = "Number of Assessment") +
  theme_bw(15) +
  theme(panel.grid = element_blank(), axis.title = element_text(face = "bold"))

ggplot(df1.2) +
  aes(x = OSP_B) +
  geom_bar(fill = "grey80") + 
  labs(x = "are there management measures implemented ?", y = "Number of Assessment") +
  theme_bw(15) +
  theme(panel.grid = element_blank(), axis.title = element_text(face = "bold"))

ggplot(df1.2) +
  aes(x = OSP_C) +
  geom_bar(fill = "grey80") + 
  labs(x = "Is there monitoring in place ?", y = "Number of Assessment") +
  theme_bw(15) +
  theme(panel.grid = element_blank(), axis.title = element_text(face = "bold"))
ggplot(df1.2) +
  aes(x = OSP_D) +
  geom_bar(fill = "grey80") + 
  labs(x = "is the MPA moving towards its conservation objectives ? ", y = "Number of Assessment") +
  theme_bw(15) +
  theme(panel.grid = element_blank(), axis.title = element_text(face = "bold"))
```
